Adaptive Approximation for Multivariate Linear Problems with Inputs Lying in a Cone

### 2019-08-17 DingHickKritzMakRICAMChapVerD.tex
Updates from Overleaf

We study the approximation of a linear solution operator $\SOL:\calF \to \calG$, which maps $f\in\calF$ to some $g\in\calG$, where we assume that $\calF$ and and $\calG$ are Banach spaces with norms $\norm[\calF]{\cdot}$ and $\norm[\calG]{\cdot}$, respectively. Our goal is to define a suitable set of input functions $\calC$, and an algorithm $\ALG(\cdot,\cdot):\calC \times (0,\infty) \to \calG$ that satisfies the error criterion:
\begin{equation} \label{DHKM:err_crit}
    \norm[\calG]{\SOL(f) - \ALG(f,\varepsilon)} \le \varepsilon \qquad \forall \varepsilon > 0, \ f \in \calC,
\end{equation}
where  $\varepsilon$ denotes the error tolerance. Such an algorithm will  contain a rule for choosing the number of series coefficients required, $n(f,\varepsilon)$, which depends on $f$ and $\varepsilon$.  The objectives of this chapter are to 
\begin{itemize}
    \item construct such a rule, 
    \item choose a set $\calC$ of input functions for which the rule is valid,  
    \item characterize the information cost of $\ALG$, 
    \item determine whether $\ALG$ has optimal information cost, and 
    \item understand the dependence of this cost on the number of input variables, $d$, as well as the error tolerance, $\varepsilon$.
\end{itemize}

@@

\DHKMchange{Let's consider an alternative cone, i.e.,
\begin{equation} \label{DHKM:altconedef}
    \calC'_{\boldsymbol{\lambda}, \boldsymbol{\omega}, A'}=\{ f \in \calF : \norm[\calF]{f} \le A'\norm[\calF']{f}\},
\end{equation}
where
$\norm[\calF']{f} =\norm[\rho]{\left( \frac{\hf(\bsk)\omega_{\bsk}}{\lambda_{\bsk}} \right)_{\bsk \in \mathbb{K}}}$ and $\omega_{\bsk}$ decreases. The norm $\norm[\calF']{\cdot}$ is weaker than the norm $\norm[\calF]{\cdot}$, so this cone is defined as the subset of $\calF$ for which the weaker norm of the function  provides an upper bound the stronger norm of the function.  The following proposition describes the relationship between the cones defined in \eqref{DHKM:pilot_cone} and \eqref{DHKM:altconedef}.
\begin{proposition}
If there exists $A,A',A'', n_1,$ and $\boldsymbol{\omega}$ such that
\[\omega_{n_1} A'' \ge A \ge \omega_1A'\left[\frac{1 + (\omega_{n_1+1}/\omega_1)^\rho}{1 - (\omega_{n_1+1}A')^\rho}\right]^{1/\rho} \text{ and} \qquad  \omega_{n_1+1}A' < 1, \]
then 
\[\calC'_{\boldsymbol{\lambda}, \boldsymbol{\omega}, A''}
\subseteq 
\calC_{\boldsymbol{\lambda}, A}
\subseteq
\calC'_{\boldsymbol{\lambda}, \boldsymbol{\omega}, A'}.\]
\end{proposition}

\begin{proof}
Let's prove $\calC'_{\boldsymbol{\lambda}, \boldsymbol{\omega}, A''}
\subseteq 
\calC_{\boldsymbol{\lambda}, A}$ at first.
Suppose $f \in \calC_{\boldsymbol{\lambda}, A}$, then by definition, we have
\begin{align*}
    \norm[\calF]{f} &\le A \norm[\rho]{\left( \frac{\hf(\bsk)}{\lambda_{\bsk}} \right)_{\bsk = 1} ^{ n_1}}
     \le A \omega_{n_1}^{-1} \norm[\rho]{\left( \frac{\hf(\bsk)\omega_{\bsk}}{\lambda_{\bsk}} \right)_{\bsk =1} ^{n_1}}
     \le A \omega_{n_1}^{-1} \norm[\calF']{f}.
\end{align*}
Thus, if $A'' \ge A\omega_{n_1}^{-1}$, i.e. $\omega_{n_1} A'' \ge A $,
we obtain $\calC'_{\boldsymbol{\lambda}, \boldsymbol{\omega}, A''}$.

Next, we show the rest part.
Suppose, $f \in \calC'_{\boldsymbol{\lambda},\boldsymbol{\omega},  A'}$, we have
\begin{align*}
    \norm[\calF']{f}^\rho- \norm[\rho]{\biggl ( \frac{\hf(\bsk_i) \omega_i }{\lambda_{\bsk_i}} \biggr) _{i=1}^{n_1}}^\rho 
	& = \norm[\rho]{\biggl ( \frac{\hf(\bsk_i) \omega_i }{\lambda_{\bsk_i}} \biggr) _{i=n_1 + 1}^{\infty}}^\rho 
     \le \omega_{n_1+1}^\rho \norm[\rho]{\biggl ( \frac{\hf(\bsk_i) }{\lambda_{\bsk_i}} \biggr) _{i=n_1 + 1}^{\infty}}^\rho \\
     & = \omega_{n_1+1}^\rho \left[ \norm[\calF]{f}^\rho - \norm[\rho]{\biggl ( \frac{\hf(\bsk_i) }{\lambda_{\bsk_i}} \biggr) _{i=1}^{n_1}}^\rho \right ]\\	
     & \le \omega_{n_1+1}^\rho \left[ A'{}^\rho \norm[\calF']{f}^\rho - \omega_1^{-\rho} \norm[\rho]{\biggl ( \frac{\hf(\bsk_i) \omega_i}{\lambda_{\bsk_i}} \biggr) _{i=1}^{n_1}}^\rho \right ].
\end{align*}
Rearrange the terms, we obtain
\begin{align*}
\norm[\calF']{f}^\rho & \le \frac{1 + (\omega_{n_1+1}/\omega_1)^\rho}{1 - (\omega_{n_1+1}A')^\rho} \norm[\rho]{\biggl ( \frac{\hf(\bsk_i) \omega_i }{\lambda_{\bsk_i}} \biggr) _{i=1}^{n_1}}^\rho \\
 \iff	\norm[\calF]{f} & \le \omega_1A'\left[\frac{1 + (\omega_{n_1+1}/\omega_1)^\rho}{1 - (\omega_{n_1+1}A')^\rho}\right]^{1/\rho} \norm[\rho]{\biggl ( \frac{\hf(\bsk_i)}{\lambda_{\bsk_i}} \biggr) _{i=1}^{n_1}}.
\end{align*}
Hence, if $A \ge \omega_1A'\left[\frac{1 + (\omega_{n_1+1}/\omega_1)^\rho}{1 - (\omega_{n_1+1}A')^\rho}\right]^{1/\rho} \text{ and } \omega_{n_1+1}A' < 1,$
we obtain $\calC_{\boldsymbol{\lambda}, A}
\subseteq
\calC'_{\boldsymbol{\lambda}, \boldsymbol{\omega}, A'}.$
\end{proof}
}

@@

\begin{lemma}\label{DHKM:lambdineq}
Consider the ordering weights,
$\lambda_{\bsk_{1}} \ge \lambda_{\bsk_{2}} \ge \cdots > 0,$
we have for $p>0$
\[ \sum_{i=1}^\infty \lambda_{d,\bsk_{i}}^p \le \frac{1}{n^{1/p}}
\left(\sum_{\bsk \in \NN_0^d}^\infty \lambda_{\bsk}^p\right)^{1/p}\]
\end{lemma}
\begin{proof}
As $\lambda_{\bsk_{1}} \ge \lambda_{\bsk_{2}} \ge \cdots > 0$, we obtain for $p>0$
\begin{align*}
    n \lambda_{\bsk_{n+1}}^p & \le 
    \lambda_{\bsk_{1}}^p + \lambda_{\bsk_{2}}^p + \cdots 
    + \lambda_{\bsk_{n}}^p  \le \sum_{\bsk \in \NN_0^d}^\infty \lambda_{\bsk}^p \\
    \Longleftrightarrow 
    \lambda_{\bsk_{n+1}}&  \le \frac{1}{n^{1/p}}
\left(\sum_{\bsk \in \NN_0^d}^\infty \lambda_{\bsk}^p \right)^{1/p}
\end{align*}
\end{proof}

>>>>

We study the approximation of a linear solution operator $\SOL:\calF \to \calG$, which maps $f\in\calF$ to some $g\in\calG$, where we assume that $\calF$ and $\calG$ are Banach spaces with norms $\norm[\calF]{\cdot}$ and $\norm[\calG]{\cdot}$, respectively. Our goal is to define a suitable set of input functions $\calC$, and an algorithm $\ALG(\cdot,\cdot):\calC \times (0,\infty) \to \calG$ that satisfies the error criterion:
\begin{equation} \label{DHKM:err_crit}
    \norm[\calG]{\SOL(f) - \ALG(f,\varepsilon)} \le \varepsilon \qquad \forall \varepsilon > 0, \ f \in \calC,
\end{equation}
where  $\varepsilon$ denotes the error tolerance. Such an algorithm will  contain a rule for choosing the number of series coefficients required, $n(f,\varepsilon)$, which depends on $f$ and $\varepsilon$.  The objectives of this chapter are to 
\begin{itemize}
    \item construct such a rule, 
    \item choose a set $\calC$ of input functions for which the rule is valid,  
    \item characterize the information cost of $\ALG$, 
    \item determine whether $\ALG$ has optimal information cost, and 
    \item understand the dependence of this cost on the number of input variables, $d$, as well as the error tolerance, $\varepsilon$.
\end{itemize}

@@

\DHKMchange{Let us consider an alternative cone, i.e.,
\begin{equation} \label{DHKM:altconedef}
    \calC'_{\boldsymbol{\lambda}, \boldsymbol{\omega}, A'}=\{ f \in \calF : \norm[\calF]{f} \le A'\norm[\calF']{f}\},
\end{equation}
where
$\norm[\calF']{f} =\norm[\rho]{\left( \frac{\hf(\bsk)\omega_{\bsk}}{\lambda_{\bsk}} \right)_{\bsk \in \mathbb{K}}}$ and $\omega_{\bsk}$ decreases. The norm $\norm[\calF']{\cdot}$ is weaker than the norm $\norm[\calF]{\cdot}$, so this cone is defined as the subset of $\calF$ for which the weaker norm of the function  provides an upper bound the stronger norm of the function.  The following proposition describes the relationship between the cones defined in \eqref{DHKM:pilot_cone} and \eqref{DHKM:altconedef}.
\begin{proposition}
If there exist $A,A',A'', n_1,$ and $\boldsymbol{\omega}$ such that
\[\omega_{n_1} A'' \ge A \ge \omega_1A'\left[\frac{1 + (\omega_{n_1+1}/\omega_1)^\rho}{1 - (\omega_{n_1+1}A')^\rho}\right]^{1/\rho} \text{ and} \qquad  \omega_{n_1+1}A' < 1, \]
then 
\[\calC'_{\boldsymbol{\lambda}, \boldsymbol{\omega}, A''}
\subseteq 
\calC_{\boldsymbol{\lambda}, A}
\subseteq
\calC'_{\boldsymbol{\lambda}, \boldsymbol{\omega}, A'}.\]
\end{proposition}

\begin{proof}
Let us prove $\calC'_{\boldsymbol{\lambda}, \boldsymbol{\omega}, A''}
\subseteq 
\calC_{\boldsymbol{\lambda}, A}$ at first.
Suppose $f \in \calC_{\boldsymbol{\lambda}, A}$, then by definition we have
\begin{align*}
    \norm[\calF]{f} &\le A \norm[\rho]{\left( \frac{\hf(\bsk)}{\lambda_{\bsk}} \right)_{\bsk = 1} ^{ n_1}}
     \le A \omega_{n_1}^{-1} \norm[\rho]{\left( \frac{\hf(\bsk)\omega_{\bsk}}{\lambda_{\bsk}} \right)_{\bsk =1} ^{n_1}}
     \le A \omega_{n_1}^{-1} \norm[\calF']{f}.
\end{align*}
Thus, if $A'' \ge A\omega_{n_1}^{-1}$, i.e. $\omega_{n_1} A'' \ge A $,
we obtain $\calC'_{\boldsymbol{\lambda}, \boldsymbol{\omega}, A''}$.

For the proof of the remaining part, 
suppose $f \in \calC'_{\boldsymbol{\lambda},\boldsymbol{\omega},  A'}$. Then we have
\begin{align*}
    \norm[\calF']{f}^\rho- \norm[\rho]{\biggl ( \frac{\hf(\bsk_i) \omega_i }{\lambda_{\bsk_i}} \biggr) _{i=1}^{n_1}}^\rho 
	& = \norm[\rho]{\biggl ( \frac{\hf(\bsk_i) \omega_i }{\lambda_{\bsk_i}} \biggr) _{i=n_1 + 1}^{\infty}}^\rho 
     \le \omega_{n_1+1}^\rho \norm[\rho]{\biggl ( \frac{\hf(\bsk_i) }{\lambda_{\bsk_i}} \biggr) _{i=n_1 + 1}^{\infty}}^\rho \\
     & = \omega_{n_1+1}^\rho \left[ \norm[\calF]{f}^\rho - \norm[\rho]{\biggl ( \frac{\hf(\bsk_i) }{\lambda_{\bsk_i}} \biggr) _{i=1}^{n_1}}^\rho \right ]\\	
     & \le \omega_{n_1+1}^\rho \left[ A'{}^\rho \norm[\calF']{f}^\rho - \omega_1^{-\rho} \norm[\rho]{\biggl ( \frac{\hf(\bsk_i) \omega_i}{\lambda_{\bsk_i}} \biggr) _{i=1}^{n_1}}^\rho \right ].
\end{align*}
Rearranging the terms, we obtain
\begin{align*}
\norm[\calF']{f}^\rho & \le \frac{1 + (\omega_{n_1+1}/\omega_1)^\rho}{1 - (\omega_{n_1+1}A')^\rho} \norm[\rho]{\biggl ( \frac{\hf(\bsk_i) \omega_i }{\lambda_{\bsk_i}} \biggr) _{i=1}^{n_1}}^\rho \\
 \iff	\norm[\calF]{f} & \le \omega_1A'\left[\frac{1 + (\omega_{n_1+1}/\omega_1)^\rho}{1 - (\omega_{n_1+1}A')^\rho}\right]^{1/\rho} \norm[\rho]{\biggl ( \frac{\hf(\bsk_i)}{\lambda_{\bsk_i}} \biggr) _{i=1}^{n_1}}.
\end{align*}
Hence, if $A \ge \omega_1A'\left[\frac{1 + (\omega_{n_1+1}/\omega_1)^\rho}{1 - (\omega_{n_1+1}A')^\rho}\right]^{1/\rho} \text{ and } \omega_{n_1+1}A' < 1,$
we obtain $\calC_{\boldsymbol{\lambda}, A}
\subseteq
\calC'_{\boldsymbol{\lambda}, \boldsymbol{\omega}, A'}.$
\end{proof}
}

@@

\begin{lemma}\label{DHKM:lambdineq}
Consider the ordered weights,
$\lambda_{\bsk_{1}} \ge \lambda_{\bsk_{2}} \ge \cdots > 0,$
then we have for $p>0$,
\[ \sum_{i=1}^\infty \lambda_{d,\bsk_{i}}^p \le \frac{1}{n^{1/p}}
\left(\sum_{\bsk \in \NN_0^d}^\infty \lambda_{\bsk}^p\right)^{1/p}\].
\end{lemma}
\PeterNote{It seems to me that on the left-hand side of the 
above inequality we need $\lambda_{\bsk_{n+1}}^p$, right?}
\begin{proof}
As $\lambda_{\bsk_{1}} \ge \lambda_{\bsk_{2}} \ge \cdots > 0$, we obtain for $p>0$
\begin{align*}
    n \lambda_{\bsk_{n+1}}^p & \le 
    \lambda_{\bsk_{1}}^p + \lambda_{\bsk_{2}}^p + \cdots 
    + \lambda_{\bsk_{n}}^p  \le \sum_{\bsk \in \NN_0^d}^\infty \lambda_{\bsk}^p \\
    \Longleftrightarrow 
    \lambda_{\bsk_{n+1}}&  \le \frac{1}{n^{1/p}}
\left(\sum_{\bsk \in \NN_0^d}^\infty \lambda_{\bsk}^p \right)^{1/p}.
\end{align*}
\end{proof}
